Models

### 2021-02-05 analysis/05_model_metrics.Rmd
Fitting with and without random slope

Fit GLM to overall metric trend
```{r}
# model for overall trend (random effect w/varying intercept)
## should be GLS according to Dornelas supplement, but glmer isn't happy
metric_mean_fit <- metrics %>%
  mutate(minyear = min(year), maxyear = max(year)) %>%
  group_nest(metric, minyear, maxyear) %>%
  mutate(model = map(data, ~lmer(logvalue ~ year + (1|study_id) + (1|rarefyID), 
                                 data = .x, control = lmerControl(optimizer ="Nelder_Mead"))))

metric_mean_coef <- metric_mean_fit %>%
  mutate(coef = map(model, tidy)) %>% 
  #mutate(confint = map(model, confint)) %>%
  select(-data, -model) %>% 
  unnest(cols = c(coef))

mean_tests <- metric_mean_fit %>%
  #filter(metric %in% c("FRic", "FEve", "FDiv", "FDis", "S")) %>%
   mutate(p_value = map(model, ~drop1(.x, test = "Chisq") %>% select(`Pr(Chi)`) %>% slice(2) %>% pull())) %>% 
  select(metric, p_value) %>%
  left_join(metric_mean_coef %>% 
              filter(term == "year") %>% 
              select(metric, estimate))
```

>>>>

Fit GLM to overall metric trend

Convergence problems!! See here:
https://joshua-nugent.github.io/allFit/
https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html
https://biologyforfun.wordpress.com/2018/04/09/help-i-have-convergence-warnings/

Maybe need something other than a gaussian distribution?

Model checks:
https://debruine.github.io/posts/normality/
performance::check_model()

```{r}
# model for overall trend (random effect w/varying intercept)
metric_mean_fit <- metrics %>%
  filter(metric %in%  c("FRic", "FEve", "FDiv", "FDis", "S")) %>%
  group_nest(metric) %>%
  mutate(model = map(data, ~lmer(logvalue ~ year + (1|study_id/rarefyID), 
                                 data = .x, control = lmerControl(optimizer ="Nelder_Mead"))),
           null_model = map(data, ~lmer(logvalue ~ (1|study_id/rarefyID), 
                                 data = .x, control = lmerControl(optimizer ="Nelder_Mead"))),
           test = map2(model, null_model, anova),
           test_res = map(test, tidy),
           coef = map(model, tidy)
           ) %>% 
  unnest(cols = c(test_res)) %>%
  select(metric, model, coef, p.value) %>%
  filter(!is.na(p.value)) %>%
  unnest(cols = c(coef)) %>%
  filter(term == "year") %>%
  select(-c(effect, group, term, statistic))

metric_mean_fit_slope <- metrics %>%
  filter(metric %in%  c("FRic", "FEve", "FDiv", "FDis", "S")) %>%
  group_nest(metric) %>%
  mutate(model = map(data, ~lmer(logvalue ~ (year|study_id/rarefyID), 
                                 data = .x, #control = lmerControl(optimizer ="Nelder_Mead"),
                                 REML = FALSE)),
           null_model = map(data, ~lmer(logvalue ~ (1|study_id/rarefyID), 
                                 data = .x, #control = lmerControl(optimizer ="Nelder_Mead"),
                                 REML = FALSE)),
           test = map2(model, null_model, anova),
           test_res = map(test, tidy),
           coef = map(model, tidy)
           ) %>% 
  unnest(cols = c(test_res)) %>%
  select(metric, model, coef, p.value) %>%
  filter(!is.na(p.value)) %>%
  unnest(cols = c(coef)) %>%
  filter(term == "year") %>%
  select(-c(effect, group, term, statistic))
```
